% !TeX root = ../main.tex
% !TeX encoding = utf8
%
% HYPERREF CONFIGURATION
% This file must be loaded AFTER all other packages
% Configures PDF metadata and cross-reference styling

% =============================================================================
% HYPERREF SETUP
% =============================================================================
\usepackage{hyperref}
\hypersetup{%
  % hidelinks,              % Uncomment for no colored links (print version)
  linkbordercolor=0.8 0 0,
  citebordercolor=0 0.8 0,
  colorlinks = true,        % Comment out or set to false for print version
  linkcolor = [rgb]{0.5, 0, 0},   % Internal links (red)
  urlcolor = [rgb]{0, 0, 0.5},    % URLs (blue)
  citecolor = [rgb]{0, 0.5, 0},   % Citations (green)
	pdftitle={\varDocumento},%
	pdfauthor={\textcopyright\ \miNombre, \miFacultad, \miUniversidad},%
  pdfsubject={\varDocumento},%
	pdfkeywords={},%
	pdfcreator={pdfLaTeX},%
}

% Alternative hypersetup (more complete)
\hypersetup{
  pdfauthor={Fernández Gutiérrez, Miguel Ángel},
  pdftitle={\varDocumento},
  pdfsubject={\varDocumento — \miGrado — Curso \miCurso},
  pdfkeywords={TFM, ciencia de datos},
  pdfproducer=mianfg,
  pdfcreator=pdflatex,
  breaklinks=true
}

% =============================================================================
% CLEVEREF CONFIGURATION (Spanish names)
% =============================================================================
\crefname{section}{sección}{secciones}
\Crefname{section}{sec.}{secs.}
\crefname{chapter}{capítulo}{capítulos}
\Crefname{chapter}{cap.}{caps.}
\crefname{part}{parte}{partes}
\Crefname{part}{parte}{partes}

\crefname{programa}{programa}{programas}
\Crefname{programa}{Programa}{Programas}
\crefname{problema}{problema}{problemas}
\Crefname{problema}{Problema}{Problemas}
\crefname{ejemplo}{ejemplo}{ejemplos}
\Crefname{ejemplo}{Ejemplo}{Ejemplos}
\crefname{proposicion}{proposición}{proposiciones}
\Crefname{proposicion}{prop.}{props.}
\crefname{lema}{lema}{lemas}
\Crefname{lema}{Lema}{Lemas}
\crefname{corolario}{corolario}{corolarios}
\Crefname{corolario}{Corolario}{Corolarios}
\crefname{teorema}{teorema}{teoremas}
\Crefname{teorema}{Teorema}{Teoremas}
\crefname{programa}{programa}{programas}
\Crefname{programa}{Programa}{Programas}
\crefname{definicion}{definición}{definiciones}
\Crefname{definicion}{Definición}{Definiciones}
\crefname{figure}{figura}{figuras}
\Crefname{figure}{Figura}{Figuras}
\crefname{tabla}{tabla}{tablas}
\Crefname{tabla}{Tabla}{Tablas}
\crefname{appendix}{apéndice}{apéndices}
\Crefname{appendix}{Apéndice}{Apéndices}
\crefname{listing}{programa}{programas}
\Crefname{listing}{Programa}{Programas}
\crefname{line}{línea}{líneas}
\Crefname{line}{Línea}{Líneas}
\crefname{page}{página}{páginas}
\Crefname{page}{pág.}{págs.}
\crefname{sistemalogico}{sistema lógico}{sistemas lógicos}
\Crefname{sistemalogico}{Sistema Lógico}{Sistemas Lógicos}
\crefname{sistemaformal}{sistema formal}{sistemas formales}
\Crefname{sistemaformal}{Sistema Formal}{Sistemas Formales}

% Conjunctions for multiple references
\newcommand{\crefpairconjunction}{ y }
\newcommand{\crefrangeconjunction}{ a }

